\TODO{What's the title?}

### 2018-09-11 doc/Paper.tex
Update PDF to 61f26b5a7e2e62a847a71aa3461d2659dad37613

\usepackage{amsmath}
\usepackage{color}
\usepackage{comment}
\usepackage{float}
\usepackage{graphicx}
\usepackage{url}
\usepackage{subcaption}
\usepackage{wrapfig}

@@

%% ODER: format ==         = "\mathrel{==}"
%% ODER: format /=         = "\neq "
%
%
\makeatletter
\@ifundefined{lhs2tex.lhs2tex.sty.read}%
  {\@namedef{lhs2tex.lhs2tex.sty.read}{}%
   \newcommand\SkipToFmtEnd{}%
   \newcommand\EndFmtInput{}%
   \long\def\SkipToFmtEnd#1\EndFmtInput{}%
  }\SkipToFmtEnd

@@

\newcommand{\Sp}{\hskip.33334em\relax}

\newcommand{\Conid}[1]{\mathit{#1}}
\newcommand{\Varid}[1]{\mathit{#1}}
\newcommand{\anonymous}{\kern0.06em \vbox{\hrule\@width.5em}}
\newcommand{\plus}{\mathbin{+\!\!\!+}}
\newcommand{\bind}{\mathbin{>\!\!\!>\mkern-6.7mu=}}
\newcommand{\rbind}{\mathbin{=\mkern-6.7mu<\!\!\!<}}% suggested by Neil Mitchell
\newcommand{\sequ}{\mathbin{>\!\!\!>}}
\renewcommand{\leq}{\leqslant}
\renewcommand{\geq}{\geqslant}
\usepackage{polytable}

%mathindent has to be defined
\@ifundefined{mathindent}%
  {\newdimen\mathindent\mathindent\leftmargini}%
  {}%

\def\resethooks{%
  \global\let\SaveRestoreHook\empty
  \global\let\ColumnHook\empty}
\newcommand*{\savecolumns}[1][default]%
  {\g@addto@macro\SaveRestoreHook{\savecolumns[#1]}}
\newcommand*{\restorecolumns}[1][default]%
  {\g@addto@macro\SaveRestoreHook{\restorecolumns[#1]}}
\newcommand*{\aligncolumn}[2]%
  {\g@addto@macro\ColumnHook{\column{#1}{#2}}}

\resethooks

\newcommand{\onelinecommentchars}{\quad-{}- }
\newcommand{\commentbeginchars}{\enskip\{-}
\newcommand{\commentendchars}{-\}\enskip}

\newcommand{\visiblecomments}{%
  \let\onelinecomment=\onelinecommentchars
  \let\commentbegin=\commentbeginchars
  \let\commentend=\commentendchars}

\newcommand{\invisiblecomments}{%
  \let\onelinecomment=\empty
  \let\commentbegin=\empty
  \let\commentend=\empty}

\visiblecomments

\newlength{\blanklineskip}
\setlength{\blanklineskip}{0.66084ex}

\newcommand{\hsindent}[1]{\quad}% default is fixed indentation
\let\hspre\empty
\let\hspost\empty
\newcommand{\NB}{\textbf{NB}}
\newcommand{\Todo}[1]{$\langle$\textbf{To do:}~#1$\rangle$}

@@

@@

@@

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[3]{}\Varid{main}\mathbin{::}\Conid{IO}\;(){}\<[E]%
\\
\>[3]{}\Varid{main}\mathrel{=}\Varid{putStrLn}\;\text{\tt \char34 Hello~world!\char34}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Functional Zippers}
  The zipper data structure was originally conceived by
  Huet\cite{huet1997zipper} to solve the problem of representing a tree together
  with a subtree that is the focus of attention, where that focus may move left,
  right, up or down the tree. Bla-bla-bla...

  Application to binary trees...
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[3]{}\mathbf{data}\;\Conid{Tree}\;\Varid{a}{}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}\mathrel{=}\Conid{Fork}\;(\Conid{Tree}\;\Varid{a})\;(\Conid{Tree}\;\Varid{a}){}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}\mid \Conid{Leaf}\,!\Varid{a}{}\<[E]%
\\[\blanklineskip]%
\>[3]{}\mathbf{data}\;\Conid{Path}\;\Varid{a}{}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}\mathrel{=}\Conid{Top}{}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}\mid \Conid{Left}\,!(\Conid{Path}\;\Varid{a})\;(\Conid{Tree}\;\Varid{a}){}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}\mid \Conid{TreeRight}\;(\Conid{Tree}\;\Varid{a})\,!(\Conid{Path}\;\Varid{a}){}\<[E]%
\\[\blanklineskip]%
\>[3]{}\mathbf{data}\;\Conid{Zipper}\;\Varid{a}\mathrel{=}\Conid{Zipper}\,!(\Conid{Path}\;\Varid{a})\;(\Conid{Tree}\;\Varid{a}){}\<[E]%
\ColumnHook
\end{hscode}\resethooks

  Application to lists...
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[3]{}\mathbf{data}\;\Conid{Path}\;\Varid{a}\mathrel{=}\Conid{Path}\;[\mskip1.5mu \Varid{a}\mskip1.5mu]\;[\mskip1.5mu \Varid{a}\mskip1.5mu]{}\<[E]%
\\[\blanklineskip]%
\>[3]{}\mathbf{data}\;\Conid{Zipper}\;\Varid{a}\mathrel{=}\Conid{Zipper}\,!(\Conid{Path}\;\Varid{a})\;[\mskip1.5mu \Varid{a}\mskip1.5mu]{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

  Generic zipper...

  An application of generic zipper that we will consider is embedding of
  attribute grammars.

\section{Attribute Grammars}
  What attribute grammars are...

  Repmin as two traversals...

  Repmin as a circular program...

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{References}
  % As per the LLNCS guidelines
  \bibliographystyle{splncs04}
  \bibliography{References}

>>>>

\usepackage{amsmath}
\usepackage{calc}
\usepackage{color}
\usepackage{comment}
\usepackage{float}
\usepackage{graphicx}
\usepackage{url}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage{pbox}
\usepackage{tikz}
\usetikzlibrary{arrows,arrows.meta}

@@

%
%
\makeatletter
\@ifundefined{lhs2tex.lhs2tex.sty.read}%
  {\@namedef{lhs2tex.lhs2tex.sty.read}{}%
   \newcommand\SkipToFmtEnd{}%
   \newcommand\EndFmtInput{}%
   \long\def\SkipToFmtEnd#1\EndFmtInput{}%
  }\SkipToFmtEnd

@@

\newcommand{\Sp}{\hskip.33334em\relax}
\newcommand{\NB}{\textbf{NB}}
\newcommand{\Todo}[1]{$\langle$\textbf{To do:}~#1$\rangle$}

@@

\newcommand{\WithMath}[2]{{\parbox{\widthof{#1}}{\centering$#2$}}}

@@

\newcommand{\mimplies}{%
  \hspace{2pt}\makebox[10pt][l]{%
    \tikz[transform canvas={scale=0.5}]{%
    \draw [double distance=3pt,
           line width=0.65pt,
           arrows={Implies[reversed]}-{Implies}] (0,0.17) -- (0.7,0.17);%
    }%
  }%
  \hspace{2pt}%
}

@@

$\mimplies\Rightarrow\implies$

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Functional Zippers}
  Zipper is a data structure commonly used in functional programming for
  traversal with fast local updates. The zipper data structure was originally
  conceived by Huet\cite{huet1997zipper} in the context of trees. We will,
  however, first consider a simpler problem: a bidirectional list traversal.

  Suppose that we would like to update a list at a specific position:
\begin{hscode}\SaveRestoreHook
~~modify~::~(a~\WithMath{->}{\rightarrow}~[a])~\WithMath{->}{\rightarrow}~{\itshape Int}~\WithMath{->}{\rightarrow}~[a]~\WithMath{->}{\rightarrow}~[a]
\\
\texfamily ~~modify~f~i~xs~=~helper~[]~xs~0
\\
\texfamily ~~~{\bfseries where}~helper~before~(x~:~after)~!j
\\
\texfamily ~~~~~~~~~~~|~j~\WithMath{==}{\equiv}~i~~~~=~before~\WithMath{++}{+\!\!+}~f~x~\WithMath{++}{+\!\!+}~after
\\
\texfamily ~~~~~~~~~~~|~\textit{otherwise}~=~helper~(before~\WithMath{++}{+\!\!+}~[x])~after~(j~+~1)
\\
\texfamily ~~~~~~~~~helper~\char95 ~[]~\char95 ~=~\textit{error}~\char34 Index~out~of~bounds.\char34 
\ColumnHook
\end{hscode}\resethooks
  Here \text{\tt modify} takes an update action \text{\tt f}\footnote{\text{\tt f} returns a list rather
  than a single element to prevent curious readers from suggesting to use a
  boxed array instead of a list.}, an index \text{\tt i}, and a list \text{\tt xs} and returns a
  new list with the \text{\tt i}'th element replaced with the result of \text{\tt f}.%
  This function "unpacks" a list, modifies one element, and "packs" the result
  into a list. If we do a lot of updates, we end up unpacking and packing the
  list over and over again -- very time-consuming for long lists. Explicitly
  working with the unpacked representation is bug-prone. A list zipper
  simplifies this.

  A zipper consists of a focus (alternatively called a hole) and surrounding
  context:
\begin{hscode}\SaveRestoreHook
~~{\bfseries data}~{\itshape Zipper}~a~=~{\itshape Zipper}~\char123 ~\char95 hole~::~a,~\char95 cxt~::~!({\itshape Context}~a)~\char125 
\\
\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
\\
\texfamily ~~{\bfseries data}~{\itshape Context}~a~=~{\itshape Context}~[a]~[a]
\\
\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
\ColumnHook
\end{hscode}\resethooks
  where the \text{\tt ListContext} keeps track of elements to the left and to the right
  of the focus. We can now define movements:
\begin{hscode}\SaveRestoreHook
~~left~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~left~({\itshape Zipper}~\char95 ~~~~({\itshape Context}~[]~\char95 ))~~~~~~~~=~{\itshape Nothing}
\\
\texfamily ~~left~({\itshape Zipper}~hole~({\itshape Context}~(l~:~ls)~rs))~=~{\itshape Just}~\$~{\itshape Zipper}~l~\$~{\itshape Context}~ls~(hole~:~rs)
\\
\texfamily 
\\
\texfamily ~~right~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~right~({\itshape Zipper}~\char95 ~~~~({\itshape Context}~\char95 ~[]))~~~~~~~~=~{\itshape Nothing}
\\
\texfamily ~~right~({\itshape Zipper}~hole~({\itshape Context}~ls~(r~:~rs)))~=~{\itshape Just}~\$~{\itshape Zipper}~r~\$~{\itshape Context}~(hole~:~ls)~rs
\ColumnHook
\end{hscode}\resethooks
  and functions for entering and leaving the zipper:
\begin{hscode}\SaveRestoreHook
~~lzEnter~::~[a]~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~lzEnter~[]~~~~~~~=~{\itshape Nothing}
\\
\texfamily ~~lzEnter~(x~:~xs)~=~{\itshape Just}~\$~{\itshape Zipper}~x~({\itshape Context}~[]~xs)
\\
\texfamily 
\\
\texfamily ~~lzLeave~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~[a]
\\
\texfamily ~~lzLeave~({\itshape Zipper}~hole~({\itshape Context}~ls~rs))~=~reverse~ls~\WithMath{++}{+\!\!+}~hole~:~rs
\ColumnHook
\end{hscode}\resethooks

  Finally, we define a local version of our \text{\tt modify} function (\TODO{Boy, is
  this function ugly...})
\begin{hscode}\SaveRestoreHook
~~lzModify~::~(a~\WithMath{->}{\rightarrow}~[a])~\WithMath{->}{\rightarrow}~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~lzModify~f~({\itshape Zipper}~hole~({\itshape Context}~ls~rs))~=~{\bfseries case}~f~hole~{\bfseries of}
\\
\texfamily ~~~~(x~:~xs)~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~x~({\itshape Context}~ls~(xs~\WithMath{++}{+\!\!+}~rs))
\\
\texfamily ~~~~[]~~~~~~~\WithMath{->}{\rightarrow}~{\bfseries case}~rs~{\bfseries of}
\\
\texfamily ~~~~~~(r~:~rs')~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~r~({\itshape Context}~ls~rs')
\\
\texfamily ~~~~~~[]~~~~~~~~\WithMath{->}{\rightarrow}~{\bfseries case}~ls~{\bfseries of}
\\
\texfamily ~~~~~~~~(l~:~ls')~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~l~({\itshape Context}~ls'~rs)
\\
\texfamily ~~~~~~~~[]~~~~~~~~\WithMath{->}{\rightarrow}~{\itshape Nothing}
\ColumnHook
\end{hscode}\resethooks

  \newpage
  using which we can perform multiple updates efficiently and with minimal code
  bloat\footnote{%
  Operator \mimplies comes from \text{\tt Control\char46{}Monad} module in \text{\tt base}
  and has the following signature:
\begin{hscode}\SaveRestoreHook
~~(\mimplies)~::~{\itshape Monad}~m~=>~(a~\WithMath{->}{\rightarrow}~m~b)~\WithMath{->}{\rightarrow}~(b~\WithMath{->}{\rightarrow}~m~c)~\WithMath{->}{\rightarrow}~a~\WithMath{->}{\rightarrow}~m~c
\ColumnHook
\end{hscode}\resethooks
  }:
\begin{hscode}\SaveRestoreHook
~~modifyExample~::~{\itshape IO}~()
\\
\texfamily ~~modifyExample~=~print~\$
\\
\texfamily ~~~~lzEnter~@{\itshape Int}~\mimplies~right
\\
\texfamily ~~~~~~~~~~~~~~~~~\mimplies~right
\\
\texfamily ~~~~~~~~~~~~~~~~~\mimplies~lzModify~(const~[])
\\
\texfamily ~~~~~~~~~~~~~~~~~\mimplies~lzModify~(return~\WithMath{.}{\circ}~(+1))
\\
\texfamily ~~~~~~~~~~~~~~~~~\mimplies~left
\\
\texfamily ~~~~~~~~~~~~~~~~~\mimplies~lzModify~(return~\WithMath{.}{\circ}~negate)
\\
\texfamily ~~~~~~~~~~~~~~~~~\mimplies~return~\WithMath{.}{\circ}~lzLeave~\$
\\
\texfamily ~~~~~~[1,~2,~3,~4,~5]
\ColumnHook
\end{hscode}\resethooks

  Consider now a binary tree data structure:
\begin{hscode}\SaveRestoreHook
~~{\bfseries data}~{\itshape Tree}~a
\\
\texfamily ~~~~=~{\itshape Fork}~({\itshape Tree}~a)~({\itshape Tree}~a)
\\
\texfamily ~~~~|~{\itshape Leaf}~!a
\\
\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
\ColumnHook
\end{hscode}\resethooks
  A binary tree zipper is slightly more insteresting than the list zipper,
  because we can move up and down the tree as well. The zipper again consists of
  a hole (a subtree we are focused on) and its surrounding context (a path from
  the hole to the root of the tree):
\begin{hscode}\SaveRestoreHook
~~{\bfseries data}~{\itshape Zipper}~a~=~{\itshape Zipper}~\char123 ~\char95 hole~::~({\itshape Tree}~a),~\char95 cxt~::~!({\itshape Context}~a)~\char125 
\\
\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
\\
\texfamily 
\\
\texfamily ~~{\bfseries data}~{\itshape Context}~a
\\
\texfamily ~~~~=~{\itshape Top}
\\
\texfamily ~~~~|~{\itshape Left}~!({\itshape Context}~a)~({\itshape Tree}~a)
\\
\texfamily ~~~~|~{\itshape TreeRight}~({\itshape Tree}~a)~!({\itshape Context}~a)
\\
\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
\ColumnHook
\end{hscode}\resethooks
  To move the zipper down, we "unpack" the current hole:
\begin{hscode}\SaveRestoreHook
~~down~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~down~({\itshape Zipper}~({\itshape Leaf}~\char95 )~\char95 )~=~{\itshape Nothing}
\\
\texfamily ~~down~({\itshape Zipper}~({\itshape Fork}~l~r)~cxt)~=~{\itshape Just}~\$~{\itshape Zipper}~r~({\itshape TreeRight}~l~cxt)
\ColumnHook
\end{hscode}\resethooks
  \text{\tt TreeContext} stores everything we need to reconstruct the hole, and \text{\tt tzUp} does
  exactly that:
\begin{hscode}\SaveRestoreHook
~~up~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~up~({\itshape Zipper}~\char95 ~{\itshape Top})~=~{\itshape Nothing}
\\
\texfamily ~~up~({\itshape Zipper}~l~({\itshape Left}~cxt~r))~=~{\itshape Just}~\$~{\itshape Zipper}~({\itshape Fork}~l~r)~cxt
\\
\texfamily ~~up~({\itshape Zipper}~r~({\itshape TreeRight}~l~cxt))~=~{\itshape Just}~\$~{\itshape Zipper}~({\itshape Fork}~l~r)~cxt
\ColumnHook
\end{hscode}\resethooks
  Implementation of \text{\tt tzLeft}, \text{\tt tzRight}, and \text{\tt tzEnter} is very similar to the
  List zipper case and is left as an exercise for the reader. \text{\tt tzLeave} differs
  slightly in that we now move all the way up rather than left:
\begin{hscode}\SaveRestoreHook
~~leave~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Tree}~a
\\
\texfamily ~~leave~z~=~{\bfseries case}~up~z~{\bfseries of}
\\
\texfamily ~~~~{\itshape Just}~z'~\WithMath{->}{\rightarrow}~leave~z'
\\
\texfamily ~~~~{\itshape Nothing}~\WithMath{->}{\rightarrow}~\char95 hole~z
\ColumnHook
\end{hscode}\resethooks

  The list and binary tree zipper we have considered here are homogeneous
  zippers: the type of focus does not change upon zipper movement. Such zippers
  can be a very useful abstraction. For example, a well-known window manager
  XMonad\cite{xmonad} uses a rose tree zipper to track the window under focus.
  For other tasks, however, one might need to traverse heterogeneous structures.
  A zipper that can accomodate such needs is usually called a generic zipper as
  it relies only on the generic structure of Algebraic Data Types (ADTs). One
  can view an ADT as an Abstract Syntax Tree (AST) where each node is a Haskell
  constructor rather than a syntax construct.

  The generic zipper we will use is very similar to the one presented
  in\cite{adams2010syz}. The most common technique in Haskell for supporting
  heterogeneous types is Existential Quantification. However, not every type can
  act as a hole. To support moving down the tree, we need the hole to be
  \textit{dissectible}, i.e. we would like to be able to dissect the value into
  the constructor and its arguments. Even though \text{\tt Data\char46{}Data\char46{}gfold} allows us to achieve
  this, we define out own typeclass which additionally allows us to propagate
  down arbitrary constraints:
\begin{hscode}\SaveRestoreHook
~~{\bfseries class}~{\itshape Dissectible}~(c~::~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~(a~::~{\itshape Type})~{\bfseries where}
\\
\texfamily ~~~~dissect~::~a~\WithMath{->}{\rightarrow}~{\itshape Left}~c~a
\\
\texfamily 
\\
\texfamily ~~{\bfseries data}~{\itshape Left}~c~expects~{\bfseries where}
\\
\texfamily ~~~~{\itshape LOne}~~::~b~\WithMath{->}{\rightarrow}~{\itshape Left}~c~b
\\
\texfamily ~~~~{\itshape LCons}~::~(c~b,~{\itshape Dissectible}~c~b)~=>~{\itshape Left}~c~(b~\WithMath{->}{\rightarrow}~expects)~\WithMath{->}{\rightarrow}~b~\WithMath{->}{\rightarrow}~{\itshape Left}~c~expects
\ColumnHook
\end{hscode}\resethooks
  For example, here is how we can make \text{\tt Tree} an instance of \text{\tt Disssectible}
\begin{hscode}\SaveRestoreHook
~~{\bfseries instance}~c~({\itshape Tree}~a)~=>~{\itshape Dissectible}~c~({\itshape Tree}~a)~{\bfseries where}
\\
\texfamily ~~~~dissect~({\itshape Fork}~l~r)~=~{\itshape LOne}~{\itshape Fork}~`{\itshape LCons}`~l~`{\itshape LCons}`~r
\\
\texfamily ~~~~dissect~x~=~{\itshape LOne}~x
\ColumnHook
\end{hscode}\resethooks
  We can unpack a \text{\tt Fork} and the zipper will thus be able to go down. \text{\tt Leaf}s,
  however, are left untouched and trying to go down from a \text{\tt Leaf} will return
  \text{\tt Nothing}.

  To allow the zipper to move left and right, we need a means to encode
  arguments to the right of the hole. Following Adams et al, we define a GADT
  representing constructor arguments to the right of the hole:
\begin{hscode}\SaveRestoreHook
~~{\bfseries data}~{\itshape Right}~c~provides~r~{\bfseries where}
\\
\texfamily ~~~~{\itshape RNil}~~::~{\itshape Right}~c~r~r
\\
\texfamily ~~~~{\itshape RCons}~::~(c~b,~{\itshape Dissectible}~c~b)~=>~b~\WithMath{->}{\rightarrow}~{\itshape Right}~c~provides~r~\WithMath{->}{\rightarrow}~{\itshape Right}~c~(b~\WithMath{->}{\rightarrow}~provides)~r
\ColumnHook
\end{hscode}\resethooks
  For example, for a tuple \text{\tt \char40{}Int\char44{}~Int\char44{}~Int\char44{}~Int\char44{}~Int\char44{}~Int\char41{}}, we can have
\begin{hscode}\SaveRestoreHook
~~lefts~=~{\itshape LOne}~(,,,,,)~`{\itshape LCons}`~1~`{\itshape LCons}`~2~`{\itshape LCons}`~3
\\
\texfamily ~~hole~=~4
\\
\texfamily ~~rights~=~5~`{\itshape RCons}`~6~`{\itshape RCons}`~{\itshape RNil}
\ColumnHook
\end{hscode}\resethooks
  generalising this a little, we arrive at:
\begin{hscode}\SaveRestoreHook
~~{\bfseries data}~{\itshape LocalContext}~c~hole~rights~parent~=
\\
\texfamily ~~~~{\itshape LocalContext}~!({\itshape Left}~c~(hole~\WithMath{->}{\rightarrow}~rights))~!({\itshape Right}~c~rights~parent)
\\
\texfamily 
\\
\texfamily ~~{\bfseries data}~{\itshape Context}~::~({\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~\WithMath{->}{\rightarrow}~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Type}~{\bfseries where}
\\
\texfamily ~~~~{\itshape RootContext}~::~\char'24~c~root.~{\itshape Context}~c~root~root
\\
\texfamily ~~~~(:>)~::~\char'24~c~parent~root~hole~rights.~(c~parent,~{\itshape Dissectible}~c~parent)
\\
\texfamily ~~~~~~~~~=>~!({\itshape Context}~c~parent~root)
\\
\texfamily ~~~~~~~~~\WithMath{->}{\rightarrow}~{\rmfamily\enskip\{-\# UNPACK  \#-\}\enskip}~!({\itshape LocalContext}~c~hole~rights~parent)
\\
\texfamily ~~~~~~~~~\WithMath{->}{\rightarrow}~{\itshape Context}~c~hole~root
\ColumnHook
\end{hscode}\resethooks
  And just like before the \text{\tt Zipper} is a product of the hole and context:
\begin{hscode}\SaveRestoreHook
~~{\bfseries data}~{\itshape Zipper}~(c~::~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~(root~::~{\itshape Type})~=
\\
\texfamily ~~~~\char'24~hole.~(c~hole,~{\itshape Dissectible}~c~hole)~=>
\\
\texfamily ~~~~~~{\itshape Zipper}~\char123 ~\char95 zHole~::~!hole
\\
\texfamily ~~~~~~~~~~~~~,~\char95 zCxt~~::~!({\itshape Context}~c~hole~root)
\\
\texfamily ~~~~~~~~~~~~~\char125 
\ColumnHook
\end{hscode}\resethooks

  Implementation of movements is quite straightforward and is left out. Please,
  refer to (\TODO{github repo}) for complete code.

  We now consider a rather interesting application of generic zipper: embedding of
  attribute grammars.

\section{Attribute Grammars}
  Attribute grammars (AGs) are an extension of context-free grammars that allow
  to specify context-sensitive syntax as well as the semantics. AGs achieve it
  by associating a set of attributes with each grammar symbol. These attributes
  are defined using evaluation rules assiciated with production rules of the
  context-free grammar.

  Attributes are then usually divided into two disjoint sets: synthesized
  attributes and the inherited attributes. Such distinction is required for the
  construction of a dependency graph. It is then used for specification of the
  evaluation order and detection of circularity. In the zipper-based embedding of
  attribute grammars we make no use of a dependency graph and thus do not divide
  attributes into classes.

  We will use repmin\footnote{%
    The repmin problem: Given a tree of integers, replace every integer with the
    minimum integer in the tree, in one pass.%
  } as a running example to illustrate how AGs can be specified using functional
  zippers. The classical solution is the following circular program:
\begin{hscode}\SaveRestoreHook
~~repmin~::~{\itshape Tree}~{\itshape Int}~\WithMath{->}{\rightarrow}~{\itshape Tree}~{\itshape Int}
\\
\texfamily ~~repmin~t~=~t'
\\
\texfamily ~~~{\bfseries where}
\\
\texfamily ~~~~(t',~m')~=~go~t~m'
\\
\texfamily ~~~~go~({\itshape Leaf}~x~~~~~~)~m~=~({\itshape Leaf}~m,~x)
\\
\texfamily ~~~~go~({\itshape Fork}~xs~ys~~)~m~=~({\itshape Fork}~xs'~ys',~min~m\char95 x~m\char95 y)
\\
\texfamily ~~~~~{\bfseries where}
\\
\texfamily ~~~~~~(xs',~m\char95 x)~=~go~xs~m
\\
\texfamily ~~~~~~(ys',~m\char95 y)~=~go~ys~m
\ColumnHook
\end{hscode}\resethooks

\begin{comment}
  Alternatively, we can formulate repmin in terms of an attribute grammar. We
  use \text{\tt localMin} attribute to represent local minimum of a subtree, \text{\tt globalMin}
  to represent the global minimum, and \text{\tt updated} for the solution for a subtree.
\end{comment}

  \TODO{This is a copy-paste} Returning to our example, in
  \cite{viera2009agsfly}, the authors identified three components for solving
  repmin: computing the minimal value, passing down the minimal value from the
  root to the leaves and constructing the resulting tree. In this section, we
  review the Attribute Grammar for repmin that was introduced by
  \cite{swierstra1998designing}, and show how each of the three components
  identified by \cite{viera2009agsfly} in that grammar can be embedded in
  Haskell using our approach.  The attribute grammar for repmin starts by
  defining the underlying data structure, i.e., binary leaf trees. The attribute
  grammar fragments presented in this section follow the standard AG notation of
  \cite{swierstra1998designing}. In this notation, we straightforwardly use the
  \text{\tt Tree~Int} datatype from \TODO{where?}.

  Having defined the structure, we need to define functionality. We start by
  reviewing the AG component that computes the minimal value of a tree:
\begin{hscode}\SaveRestoreHook
~~SYN~{\itshape Tree}~{\itshape Int}~[localMin~:~{\itshape Int}]
\\
\texfamily ~~SEM~{\itshape Tree}~|~{\itshape Leaf}~lhs.localMin~=~@value
\\
\texfamily ~~~~~~~~~~~|~{\itshape Fork}~lhs.localMin~=~@left.localMin~`min`~@right.localMin
\ColumnHook
\end{hscode}\resethooks
  This component declares, using the \text{\tt SYN} keyword, that elements of type \text{\tt Tree}
  synthesize an attribute \text{\tt localMin} of type \text{\tt Int}. Then, a \text{\tt SEM} sentence
  defines how \text{\tt localMin} is computed: when the current tree is a leaf, clearly
  its minimal value is the leaf value itself; when it is the fork of two other
  trees (the \text{\tt left} and the \text{\tt right} subtrees), we compute the minimal values of
  each subtree (i.e., their \text{\tt localMin} attribute), and then their minimal value
  (function \text{\tt min}). In this notation, \text{\tt lhs} refers to the left-hand side symbol
  of the production and @ prefixes a reference to a field.

  Having implemented the first of the three components that solve repmin, we now
  consider the remaining two. We start by implementing the construction of the
  result of repmin, a tree with all leaves being the minimum of the original
  one.
\begin{hscode}\SaveRestoreHook
~~SYN~{\itshape Tree}~{\itshape Int}~[updated~:~{\itshape Tree}~{\itshape Int}]
\\
\texfamily ~~SEM~{\itshape Tree}~|~{\itshape Leaf}~lhs.updated~=~{\itshape Leaf}~@lhs.globalMin
\\
\texfamily ~~~~~~~~~~~|~{\itshape Fork}~lhs.updated~=~{\itshape Fork}~@left.updated~@right.updated
\ColumnHook
\end{hscode}\resethooks
  The implementation of sres places in each leaf of a tree the value of the ival
  attribute. This value corresponds to the minimal value of the global tree,
  that still needs to be passed down to all the nodes in the tree. This
  corresponds exactly to the third component that we still need to implement. In
  order to bind the minimal value being computed (attribute smin) with the
  minimal value that is passed down through the tree (attribute ival), it is
  common, in the AG setting, to add a new data-type definition,

  In Haskell, these attributes are simply functions:

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% As per the LLNCS guidelines
\bibliographystyle{splncs04}
\bibliography{References}
